---
output:
  pdf_document: default
  html_document: default
---

# load packages
```{r}  
library(phyloseq)
library(ggplot2)
library(FSA)
library(vegan)

theme_set(theme_classic())
theme_set(theme_bw())

col <- c("#e7298a", "#1b9e77", "#d95f02", "#7570b3")

```

## make data objects for phyloseq (OTU table are unique taxa only)
```{r}
#OTU table (matrix with no tax information)
otu <- read.csv("OTUa.csv", head = T, row.names = 1)
#MetaData table (row names must be sample names from OTU table)
env <- read.csv("env.csv", head = T, row.names = 1)
```

## make into matrix
```{r}
otuMat <- as.matrix(otu)
```

##Convert to phyloseq objects and combine to make phyloseq object
```{r}
OTU <- otu_table(otuMat, taxa_are_rows = T)
ENV <- sample_data(env)
data <- phyloseq(OTU, ENV)
#look at what you have
data
```

#How many reads per sample
```{r}
sample_sums(data)
```

#remove sample S04, 46 and 47 as they have low numbers of reads get 45K per sample
```{r}
data = prune_samples(sample_sums(data)>=31318, data)

data1 <- subset_samples(data, Month != "Sept") # remove september samples, not used
data2 <- subset_samples(data1, salt_level != "S3R2 (BR)")

dataRare <- rarefy_even_depth(data2, rngseed = 1121983) # rerefy data to even sampling depth

sample_sums(dataRare)
```

#Alpha diveristy

# remove tank walls data
```{r}
alpha <- subset_samples(dataRare, TimePoint != "TankWalls") 

sample_sums(alpha)
```

#as a box plot
```{r}
#pdf("Figure1b.pdf", width = 5.5, height = 4)
a <- plot_richness(alpha,  x="test2", color="treatment",measures=c("Observed"))+
geom_boxplot()+  
 labs(x ="", title = "(b) Species richness", y = "Number of taxa")+ # name axis
  scale_x_discrete(labels=c("T0","T1","T2","T1","T2","T1","T2"))+
  scale_color_manual(name="Salinity", values = col, breaks=c("T0","Low","Medium","High"))+ # custom pallet and change order
  theme(axis.text.x=element_text(angle=0,hjust=.5))+
 # theme(axis.text.x = element_text(angle=0), hjust=0.95)+ # change x axis angle
  #axis.ticks = element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank())+
  ylim(0,1500)+ 
  annotate("text", x = 2.5, y = 1500, label = "Low", size = 4.5)+
  annotate("text", x = 4.5, y = 1500, label = "Medium", size = 4.5)+
  annotate("text", x = 6.5, y = 1500, label = "High", size = 4.5)+
  annotate("segment", x = 1.6, xend = 3.4, y = 1430, yend = 1430, cex = .8)+
  annotate("segment", x = 3.7, xend = 5.4, y = 1430, yend = 1430, cex = .8)+
  annotate("segment", x = 5.7, xend = 7.4, y = 1430, yend = 1430, cex = .8)

a
```

#export code
```{r}
pdf("Alpha Div.pdf", width = 6.65354331, height = 3.326771655)

postscript("Alpha Div.eps", width = 6.65354331, height = 3.326771655,  paper = "special",  horizontal=F)

dev.off()
```

#Kruskal-Wallis test 

#GET DATA FOR TEST (load CSV with diveristy data for each sample, calculated with estimate_richness() in phyloseq)
```{R}
richness <- read.csv("richA.csv", head = T)
richness
```

#effect of salt T1, with DunnTest for posthoc (richness = species richness, shannon = shannon diveirsty)
```{r}
test1 <- richness[richness$TimePoint == "T1" ,] # get time point 1 data

rich_saltLevel <- kruskal.test(Observed ~ treatment, data = test1)

rich_saltLevel

#Kruskal-Wallis chi-squared = 9.9257, df = 2, p-value = 0.006993


rich_saltLevel_Post = dunnTest(Observed ~ treatment,
              data=test1,
              method="bh")     

rich_saltLevel_Post 

#1    high - low -3.0289607 0.002453966 0.007361897
#2 high - medium -2.1921939 0.028365506 0.042548258
#3  low - medium  0.9621405 0.335979047 0.335979047
```


#effect of salt T2, with DunnTest for posthoc
```{r}
test2 <- richness[richness$TimePoint == "T2" ,] # Get T2 data only

rich_saltLevel <- kruskal.test(Observed ~ treatment, data = test2)

rich_saltLevel

#Kruskal-Wallis chi-squared = 5.9154, df = 2, p-value = 0.05194


rich_saltLevel_Post = dunnTest(Observed ~ treatment,
              data=test1,
              method="bh")     

rich_saltLevel_Post 

#    high - low -3.0289607 0.002453966 0.007361897
# high - medium -2.1921939 0.028365506 0.042548258*
#  low - medium  0.9621405 0.335979047 0.335979047
```


#Low salt by time
```{r}
test2 <- richness[richness$treatment == "low",] # get low salt data only

test2a <- richness[richness$treatment == "low" | richness$TimePoint =="T0",] # Get T0 and low salt
test2a <- test2a[test2a$TimePoint =="T0" | test2a$TimePoint =="T1",] # Remove T2

test2b <- richness[richness$treatment == "low" | richness$TimePoint =="T0",] # Get T0 and low salt
test2b <- test2b[test2a$TimePoint =="T0" | test2b$TimePoint =="T2",] # Remove T1
```


```{r}
#T1 v T2

rich_low <- kruskal.test(Observed ~ TimePoint, data = test2)

rich_low

#Kruskal-Wallis chi-squared = 6, df = 1, p-value = 0.01431

#T0 v T1

rich_lowa <- kruskal.test(Observed ~ TimePoint, data = test2a)

rich_lowa

#Kruskal-Wallis chi-squared = 4.5, df = 1, p-value = 0.03389 *

#T0 v T2

rich_lowb <- kruskal.test(Observed ~ TimePoint, data = test2b)

rich_lowb

#Kruskal-Wallis chi-squared = 4.5, df = 1, p-value = 0.03389 *
```

#med salt overtime
```{r}
test3 <- richness[richness$treatment == "medium",]

test3a <- richness[richness$treatment == "medium" | richness$TimePoint =="T0",]
test3a <- test3a[test3a$TimePoint =="T0" | test3a$TimePoint =="T1",]


test3b <- richness[richness$treatment == "medium" | richness$TimePoint =="T0",]
test3b <- test3b[test3a$TimePoint =="T0" | test3b$TimePoint =="T2",]

#T1 v T2

rich_med <- kruskal.test(Observed ~ TimePoint, data = test3)

rich_med

#Kruskal-Wallis chi-squared = 6.8182, df = 1, p-value = 0.009023



#T0 v T1

rich_lowa <- kruskal.test(Observed ~ TimePoint, data = test3a)

rich_lowa

#Kruskal-Wallis chi-squared = 5, df = 1, p-value = 0.02535


#T0 v T2

rich_lowb <- kruskal.test(Observed ~ TimePoint, data = test3b)

rich_lowb

#Kruskal-Wallis chi-squared = 5, df = 1, p-value = 0.02535
```


#High salt overtime
```{r}
test4 <- richness[richness$treatment == "high",]

test4a <- richness[richness$treatment == "high" | richness$TimePoint =="T0",]
test4a <- test4a[test4a$TimePoint =="T0" | test4a$TimePoint =="T1",]


test4b <- richness[richness$treatment == "high" | richness$TimePoint =="T0",]
test4b <- test4b[test4a$TimePoint =="T0" | test4b$TimePoint =="T2",]

#T1 v T2

rich_high <- kruskal.test(Observed ~ TimePoint, data = test4)

rich_high

#Kruskal-Wallis chi-squared = 3.75, df = 1, p-value = 0.05281


#T0 v T1

rich_lowa <- kruskal.test(Observed ~ TimePoint, data = test4a)

rich_lowa

#Kruskal-Wallis chi-squared = 5, df = 1, p-value = 0.02535

#T0 v T2

rich_lowb <- kruskal.test(Observed ~ TimePoint, data = test4b)

rich_lowb

#Kruskal-Wallis chi-squared = 1.3333, df = 1, p-value = 0.2482
```
