---
output:
  pdf_document: default
  html_document: default
---

# load packages
```{r}  
library(phyloseq)
library(ggplot2)
library(vegan)
library(DESeq2)

theme_set(theme_classic())
theme_set(theme_bw())

col <- c("#e7298a", "#1b9e77", "#d95f02", "#7570b3")

```

## make data objects for phyloseq
```{r}
#OTU table (matrix with no tax information)
otu <- read.csv("otu.csv", head = T, row.names = 1)
#Tax table (with OTU IDs that match OTU table and no duplicates)
tax <- read.csv("tax.csv", head = T, row.names = 1)
#MetaData table (row names must be sample names from OTU table)
env <- read.csv("env.csv", head = T, row.names = 1)
```

## make into matrix
```{r}
otuMat <- as.matrix(otu)
taxMat <- as.matrix(tax)
```

##Convert to phyloseq objects and combine to make phyloseq object
```{r}
OTU <- otu_table(otuMat, taxa_are_rows = T)
TAX <- tax_table(taxMat)
ENV <- sample_data(env)
data <- phyloseq(OTU, TAX, ENV)
#look at what you have
data
```

#How many reads per sample
```{r}
sample_sums(data)
```

#remove sample S04, 46 and 47 as they have low numbers (they are tank walls and day 1 which are not used)
```{r}
data = prune_samples(sample_sums(data)>=16000, data)

data1 <- subset_samples(data, Month != "Sept") # remove september samples, not used
data2 <- subset_samples(data1, salt_level != "S3R2 (BR)") # unsued duplicate

dataRare <- rarefy_even_depth(data2, rngseed = 1121983) # rerefy data to even sampling depth

sample_sums(dataRare)
```

# MDS make presance absance distance matrix for NMDS
```{R}
ord <- ordinate(dataRare, "NMDS", "jaccard", binary = TRUE) # make the distance matrix and ordination
```

```{r}
plot_ordination(dataRare, ord, type="samples", color="treatment", shape="TimePoint")+ # plots the ordination
  scale_color_manual(name="Salinity", values = col, breaks=c("T0","Low","Medium","High"))+ # custom pallet and change order
  scale_shape_discrete(name="Time Point")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  geom_point(size=3.5)
```

# Permanova to test effect of salt level and time (striped out GV and tank walls)
```{r}
salt.month <- subset_samples(dataRare, TimePoint != "T0") # just get the comparible samples
salt.month <- subset_samples(salt.month, TimePoint != "TankWalls") # rmove tank walls

distanceMatrix.salt.month <- phyloseq::distance(salt.month, method = "jaccard", binary = T) # make a distance matrix

envirodata.salt.month <- data.frame(sample_data(salt.month)) # data frame of your enviro vars
otudata.salt.month <- data.frame(otu_table(salt.month)) # data frame of your OTUs
otudata.salt.month <- t(otudata.salt.month) # traspose your data frame so you can run this data in vegan

adonis(distanceMatrix.salt.month ~ salt + TimePoint, data = envirodata.salt.month) # run the modle
```

# tank walls, are they different to algae? 
# not the one, check that algae biofild is sig dif to tank wall
# PERMANOVA tank wall v algae
# Significant difference between bacteria for tank wall v algae, but not as important as the salt level. 
```{r}
biofilm <- subset_samples(dataRare, Month == "Oct") # just get the comparible samples

distanceMatrix.biofilm <- phyloseq::distance(biofilm, method = "jaccard", binary = T) # make a distance matrix

envirodata.biofilm <- data.frame(sample_data(biofilm)) # data frame of your enviro vars
otudata.biofilm <- data.frame(otu_table(biofilm)) # data frame of your OTUs
otudata.biofilm <- t(otudata.biofilm) # traspose your data frame so you can run this data in vegan

adonis(distanceMatrix.biofilm ~ salt * origin, data = envirodata.biofilm) # run the modle
```

# differential_abundance with deseq2
# get data and calculate the model
```{r}
#get samples for comparison, not normalized

alt.month <- subset_samples(data, TimePoint != "TankWalls") # rmove tank walls
alt.month <- subset_samples(alt.month, treatment != "T0") # remove time 0
alt.month <- subset_taxa(alt.month, superkingdom!="") # filter out uncalisfyed Contigs
filtS <- subset_taxa(alt.month, superkingdom=="Bacteria") # keep bacteria
shortData <- prune_taxa(names(sort(taxa_sums(filtS),TRUE)[1:200]), filtS) # select top dominant taxa

sample_data(shortData) # check data

diff <- phyloseq_to_deseq2(shortData, ~ TimePoint + treatment) 

gm_mean <- function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans = apply(counts(diff), 1, gm_mean)
diff = estimateSizeFactors(diff, geoMeans = geoMeans)
diffRes = DESeq(diff, fitType="local")

save(diffRes, file="diff.RData") # save the data for table
```

#Results table and data to plot for selected comparison (medium v high)
```{r}
res = results(diffRes)
res = res[order(res$padj, na.last=NA), ]
alpha = 0.01 # p value = 99%
sigtab = res[(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(dataRare)[rownames(sigtab), ], "matrix"))
head(sigtab)

posigtab = sigtab[sigtab[, "log2FoldChange"] > 0, ]
posigtab = posigtab[, c("baseMean", "log2FoldChange", "lfcSE", "padj", "phylum", "class", "family", "genus")]

head <- posigtab

write.csv(sigtab, "HighVMed.csv") # save data for table
```

#Plot results
```{r}
library("ggplot2")
theme_set(theme_bw())
sigtabgen = subset(sigtab, !is.na(genus)) # set genera level

# make color pallete 

library("RColorBrewer") 
getPalette <- colorRampPalette(brewer.pal(9, "Set1"))
speciesList <-  sigtabgen$class
speciesPalette <- getPalette(length(speciesList)) # make a pallete this long
names(speciesPalette) = speciesList  # lists what the matches will be.

speciesPalette <- gsub("#CA2832","Pink", speciesPalette) # edit the species pallet to have better contrast bettween conditions.

pdf("FigureS1.pdf", width = 7, height = 5)

x = tapply(sigtabgen$log2FoldChange, sigtabgen$genus, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$genus = factor(as.character(sigtabgen$genus), levels=names(x))
# family order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$genus, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$genus = factor(as.character(sigtabgen$genus), levels=names(x))

ggplot(sigtabgen, aes(y=genus, x=log2FoldChange, color=class)) + 
  geom_point(size=3) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  scale_color_manual(values=speciesPalette)+
  annotate("text", x = -12, y = 32, label = "High salinity", size = 3)+
  annotate("text", x = +8.5, y = 32, label = "Medium salinity", size = 3)+
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5), panel.grid.major = element_blank(),  panel.grid.minor = element_blank() )

dev.off
```